---
title: "Analysis"
author: "Sonam Tenzin"
date: "2025-01-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this analysis, we will explore various factors that could be related to stroke occurrences. We will visualize data distributions using boxplots and bar plots, and then build a predictive model using Random Forest to predict the likelihood of a stroke.

## Data Preprocessing

First, let's load the dataset and perform necessary preprocessing, including checking for missing values and handling them.

```{r}
# Load dataset
data <- read.csv("filepath")

# Check for missing values
colSums(is.na(data))

# Handle missing values (e.g., fill BMI missing values with median)
data$bmi[is.na(data$bmi)] <- median(data$bmi, na.rm = TRUE)

# Convert categorical columns to factors
data$gender <- as.factor(data$gender)
data$hypertension <- as.factor(data$hypertension)
data$heart_disease <- as.factor(data$heart_disease)
data$ever_married <- as.factor(data$ever_married)
data$work_type <- as.factor(data$work_type)
data$Residence_type <- as.factor(data$Residence_type)
data$smoking_status <- as.factor(data$smoking_status)
data$stroke <- as.factor(data$stroke)  # Target variable

# Boxplot for Age by Stroke
ggplot(data, aes(x = stroke, y = age, fill = stroke)) +
  geom_boxplot() +
  labs(title = "Boxplot of Age by Stroke", x = "Stroke", y = "Age") +
  theme_minimal()
# Boxplot for Average Glucose Level by Stroke
ggplot(data, aes(x = stroke, y = avg_glucose_level, fill = stroke)) +
  geom_boxplot() +
  labs(title = "Boxplot of Avg Glucose Level by Stroke", x = "Stroke", y = "Avg Glucose Level") +
  theme_minimal()
# Boxplot for BMI by Stroke
ggplot(data, aes(x = stroke, y = bmi, fill = stroke)) +
  geom_boxplot() +
  labs(title = "Boxplot of BMI by Stroke", x = "Stroke", y = "BMI") +
  theme_minimal()
# Bar plot for Gender vs Stroke
ggplot(data, aes(x = gender, fill = stroke)) +
  geom_bar(position = "dodge") +
  labs(title = "Gender vs Stroke", x = "Gender", y = "Count") +
  theme_minimal()
# Bar plot for Hypertension vs Stroke
ggplot(data, aes(x = hypertension, fill = stroke)) +
  geom_bar(position = "dodge") +
  labs(title = "Hypertension vs Stroke", x = "Hypertension", y = "Count") +
  theme_minimal()
# Bar plot for Heart Disease vs Stroke
ggplot(data, aes(x = heart_disease, fill = stroke)) +
  geom_bar(position = "dodge") +
  labs(title = "Heart Disease vs Stroke", x = "Heart Disease", y = "Count") +
  theme_minimal()

### Model Building

# Install necessary packages
install.packages("caret")
library(caret)

# Split the data (60% train, 40% test)
set.seed(123)  # For reproducibility
trainIndex <- createDataPartition(data$stroke, p = 0.6, list = FALSE)

# Create training and testing datasets
trainData <- data[trainIndex, ]
testData <- data[-trainIndex, ]

### Training the Random Forest Model

library(randomForest)

# Train random forest model
model_rf <- randomForest(stroke ~ ., data = trainData, importance = TRUE)

# Variable importance plot
varImpPlot(model_rf)

### Model prediction and Evaluation

# Predict using the model
predictions <- predict(model_rf, testData, type = "class")

# Confusion Matrix
confusionMatrix(predictions, testData$stroke)

# Calculate Accuracy
accuracy <- sum(predictions == testData$stroke) / nrow(testData)
print(paste("Accuracy:", round(accuracy * 100, 2), "%"))


### Explanation:
1. **YAML Header**: At the top of the document, you specify the title and output format (e.g., HTML).
2. **R Code Chunks**: Code chunks are enclosed within triple backticks (` ```{r} ... ``` `). These chunks will execute the R code when knitting the document.
3. **Markdown Syntax**: Use markdown for sections like `## Introduction`, `## Data Preprocessing`, etc. to provide a clear structure to the document.
4. **Inline R Code**: For displaying dynamic results, you can embed R code like the accuracy result directly into the markdown text by using `r round(accuracy * 100, 2)`.

### How to Knit the Document:
1. Save the above content into a `.Rmd` file.
2. Open the file in RStudio.
3. Click the **Knit** button to generate an HTML (or PDF) output.

This will generate a well-organized report that contains your analysis, visualizations, and model results in an easily readable format.

